\bf Landslide: \\ Systematic Dynamic Race Detection in Kernel Space

### 2012-05-18 futurework.tex
fuzz things for dave & remove fixmes
add nadim's picture

\revision{
\renewcommand\bugnum[2]{#1 {\scriptsize \em (#2)}}
\begin{table}[t]
	\centering
	\footnotesize
	\begin{tabular}{|l|l||c|c|c|}
	\hline
	{\bf Bug} & {\bf Direction} & {\bf Time} & {\bf Backtracks} & {\bf Trace Length} \\
	\hline\hline
	\multirow{2}{*}{POBBLES vanish/vanish (a)} & Forwards & \bugnum{124.8}{4.9} & 1264 & 7 \\
	& Backwards & \bugnum{57.1}{1.7} & 376 & 11 \\
	\hline
	\multirow{2}{*}{POBBLES vanish/vanish (b)} & Forwards & \bugnum{106.3}{5.7} & 1030 & 7 \\
	& Backwards & \bugnum{51.5}{2.1} & 376 & 10 \\
	\hline
	\multirow{2}{*}{POBBLES wait/wait} & Forwards & \bugnum{62.6}{3.4} & 132 & 8 \\
	& Backwards & \bugnum{27.9}{0.8} & 17 & 13 \\
	\hline
	\multirow{2}{*}{POBBLES thread\_fork/vanish} & Forwards & \bugnum{21.4}{0.7} & 0 & 6 \\
	& Backwards & \bugnum{37.4}{1.1} & 70 & 6 \\
	\hline
	\multirow{2}{*}{LudicrOS vanish/vanish} & Forwards & \bugnum{14.7}{0.9} & 4 & 3 \\
	& Backwards & \bugnum{13.7}{0.7} & 3 & 7 \\
	\hline
	\multirow{2}{*}{LudicrOS yield/vanish} & Forwards & \bugnum{13}{0.9} & 3 & 3 \\
	& Backwards & \bugnum{11.4}{0.4} & 0 & 5 \\
	\hline
	\end{tabular}
	\caption{Comparison of forwards and backwards explorations for the six case study bugs (using decision points on \texttt{mutex\_lock} but not \texttt{mutex\_unlock}).}
	\label{fig:backwards}
\end{table}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.47\textwidth]{graphs/backwards-time.pdf}
	\hspace{0.02\textwidth}
	\includegraphics[width=0.48\textwidth]{graphs/backwards-trace.pdf}
	\caption{In general, backwards exploration finds bugs faster, but forwards exploration produces shorter (more understandable) decision traces.}
	\label{fig:backwards-graphs}
\end{figure}
}

@@

\revision{Table~\ref{fig:backwards} and Figure~\ref{fig:backwards-graphs} compare the tree explorations and resulting decision traces. The one exception is the POBBLES thread\_fork/vanish bug, which Landslide found immediately during forwards exploration, because the race actually required no forced preemptions to expose.}

@@

When using decision points on just \texttt{mutex\_lock} or \texttt{mutex\_unlock}, the explorations for each test case took similar time. However, when using decision points on both mutex functions, \texttt{double\_wait} took substantially longer, despite the tree structure suggesting they should have been comparable, with \texttt{double\_wait} even being perhaps slightly faster. Figure~\ref{fig:nadim-stats} shows a comparison of the tree anatomy for these two test cases.

\begin{table}[h]
	\small
	\centering
	\begin{tabular}{|l||c|c|}
		\hline
		& \texttt{double\_wait} & \texttt{vanish\_vanish} \\
		\hline\hline
		Exploration time & {\bf 30 minutes} & {\bf 5 minutes} \\
		\hline
		Backtracks & {\bf 1,401} & {\bf 1,511} \\
		\hline
		Decision points & 10,355 & 5,690 \\
		\hline
		Average branch depth & 18.34 & 16.00 \\
		\hline
		Average instructions/branch& 1,975,033 & 1,984,249 \\
		\hline
		Total instructions & 206,757,601 & 24,907,516 \\
		\hline
		Backtrack distance minimum & {\bf 6} & {\bf 2} \\
		\hline
		Backtrack distance maximum & 17 & 16 \\
		\hline
		Backtrack distance average & {\bf 7.38}  & {\bf 3.75} \\
		\hline
		% FIXME: add stddev?
	\end{tabular}
	\caption{Comparison of expensive \texttt{double\_wait} test and cheap \texttt{vanish\_vanish} test. The number of branches and depth of each branch are similar, but \texttt{double\_wait}'s backtracks were significantly longer, causing more total instructions to be executed.}
	\label{fig:nadim-stats}
\end{table}

Despite the two trees having nearly equal numbers of branches and instructions executed per branch, \texttt{double\_wait} took significantly more time. This is because Landslide's backtracks during \texttt{double\_wait} were significantly {\em longer} than the backtracks in \texttt{vanish\_vanish} on average; i.e., the backtracks were targetting an earlier point during the test's execution.
Hence, Landslide had to execute more total instructions overall to explore the longer branches in \texttt{double\_wait}.\footnote{
Of note, if backtracking had been implemented by restarting the test each time; i.e., executing the full depth of every branch from the root of the tree, the repeated work would have caused the execution times to be equal. The fact that Simics's backtracking avoids the cost of execution that happened before the backtrack target is to credit for this discrepancy.}
% as visualised in figure\ldots TODO

>>>>

\revision{
\renewcommand\bugnum[2]{#1 {\scriptsize \em (#2)}}
\begin{table}[t]
	\centering
	\small
	\begin{tabular}{|l|l||c|c|c|}
	\hline
	{\bf Bug} & {\bf Direction} & {\bf Time} & {\bf Backtracks} & {\bf Trace Length} \\
	\hline\hline
	\multirow{2}{*}{POBBLES vanish/vanish (a)} & Forwards & \bugnum{124.8}{4.9} & 1264 & 7 \\
	& Backwards & \bugnum{57.1}{1.7} & 376 & 11 \\
	\hline
	\multirow{2}{*}{POBBLES vanish/vanish (b)} & Forwards & \bugnum{106.3}{5.7} & 1030 & 7 \\
	& Backwards & \bugnum{51.5}{2.1} & 376 & 10 \\
	\hline
	\multirow{2}{*}{POBBLES wait/wait} & Forwards & \bugnum{62.6}{3.4} & 132 & 8 \\
	& Backwards & \bugnum{27.9}{0.8} & 17 & 13 \\
	\hline
	\multirow{2}{*}{POBBLES thread\_fork/vanish} & Forwards & \bugnum{21.4}{0.7} & 0 & 6 \\
	& Backwards & \bugnum{37.4}{1.1} & 70 & 6 \\
	\hline
	\multirow{2}{*}{LudicrOS vanish/vanish} & Forwards & \bugnum{14.7}{0.9} & 4 & 3 \\
	& Backwards & \bugnum{13.7}{0.7} & 3 & 7 \\
	\hline
	\multirow{2}{*}{LudicrOS yield/vanish} & Forwards & \bugnum{13}{0.9} & 3 & 3 \\
	& Backwards & \bugnum{11.4}{0.4} & 0 & 5 \\
	\hline
	\end{tabular}
	\caption{Comparison of forwards and backwards explorations for the six case study bugs (using decision points on \texttt{mutex\_lock} but not \texttt{mutex\_unlock}).}
	\label{fig:backwards}
\end{table}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.48\textwidth]{graphs/backwards-time.pdf}
	\hspace{0.01\textwidth}
	\includegraphics[width=0.48\textwidth]{graphs/backwards-trace.pdf}
	\caption{In general, backwards exploration finds bugs faster, but forwards exploration produces shorter (more understandable) decision traces. (Lower is better in both graphs.)}
	\label{fig:backwards-graphs}
\end{figure}
}

@@

\revision{Table~\ref{fig:backwards} and Figure~\ref{fig:backwards-graphs} compare the tree explorations and resulting decision traces. The one exception to backwards being faster is the POBBLES thread\_fork/vanish bug, which Landslide found immediately during forwards exploration, because the race actually required no forced preemptions to expose.}

@@

When using decision points on just \texttt{mutex\_lock} or \texttt{mutex\_unlock}, the explorations for each test case took similar time. However, when using decision points on both mutex functions, \texttt{double\_wait} took substantially longer, despite the tree structure suggesting they should have been comparable, with \texttt{double\_wait} even being perhaps slightly faster. Table~\ref{fig:nadim-stats} shows a comparison of the tree anatomy for these two test cases.

\begin{table}[h]
	\small
	\centering
	\begin{tabular}{|l||c|c|}
		\hline
		& \texttt{double\_wait} & \texttt{vanish\_vanish} \\
		\hline\hline
		Exploration time & {\bf 30 minutes} & {\bf 5 minutes} \\
		\hline
		Backtracks & {\bf 1,401} & {\bf 1,511} \\
		\hline
		Decision points & 10,355 & 5,690 \\
		\hline
		Average branch depth & 18.34 & 16.00 \\
		\hline
		Average instructions/branch& {\bf 1,975,033} & {\bf 1,984,249} \\
		\hline
		Total instructions & 206,757,601 & 24,907,516 \\
		\hline
		Backtrack distance minimum & {\bf 6} & {\bf 2} \\
		\hline
		Backtrack distance maximum & 17 & 16 \\
		\hline
		Backtrack distance average & {\bf 7.38}  & {\bf 3.75} \\
		\hline
	\end{tabular}
	\caption{Comparison of expensive \texttt{double\_wait} test and cheap \texttt{vanish\_vanish} test. The number of branches and depth of each branch are similar, but \texttt{double\_wait}'s backtracks were significantly longer, causing more total instructions to be executed.}
	\label{fig:nadim-stats}
\end{table}

Despite the two trees having nearly equal numbers of branches and instructions executed per branch, \texttt{double\_wait} took significantly more time. This is because Landslide's backtracks during \texttt{double\_wait} were significantly {\em longer} than the backtracks in \texttt{vanish\_vanish} on average; i.e., the backtracks were targetting an earlier point during the test's execution.
Hence, Landslide had to execute more total instructions overall to explore the longer branches in \texttt{double\_wait}.\footnote{
Of note, if backtracking had been implemented by restarting the test each time; i.e., executing the full depth of every branch from the root of the tree, the repeated work would have caused the execution times to be equal. The fact that Simics's backtracking avoids the cost of execution that happened before the backtrack target is to credit for this discrepancy.}
We show an exaggerated visualisation of this tree structure in Figure~\ref{fig:nadim}

\begin{figure}[h]
	\centering
	\includegraphics[width=0.9\textwidth]{nadim.pdf}
	\caption{Though the execution trees in \texttt{double\_wait} and in \texttt{vanish\_vanish} had the same number of branches, with each approximately as many instructions deep, the backtracks in \texttt{double\_wait} were much longer, resulting in more {\em total} instructions executed during the exploration.}
	\label{fig:nadim}
\end{figure}

### 2012-05-18 related.tex
add nadim's picture

Finally, DeMeter~\cite{demeter} is a more recent tool for systematic exploration that introduces Dynamic Interface Reduction as a strategy for constraining the size of the state space, which means targetting the testing technique on only one component of the program at a time. Landslide makes use of the same general concept in its user interface (Section~\ref{sec:using}) by providing the user with options to constrain decision point identification to certain subsets of the kernel.

>>>>

Finally, DeMeter~\cite{demeter} is a more recent tool for systematic exploration that introduces Dynamic Interface Reduction as a strategy for constraining the size of the state space, which means targetting the testing technique on only one component of the program at a time. Landslide makes use of the same general concept in its user interface (Chapter~\ref{sec:using}) by providing the user with options to constrain decision point identification to certain subsets of the kernel.
